---
title: java
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
source("_knitr_config.R")
```

*********************************************************************

## links

* https://medium.com/@brunofrascino/working-with-multiple-java-versions-in-macos-9a9c4f15615a  
* https://www.w3schools.com/java/default.asp  
* https://stackoverflow.com/questions/52524112/how-do-i-install-java-on-mac-osx-allowing-version-switching  
* http://davidcai.github.io/blog/posts/install-multiple-jdk-on-mac/  
* http://www.lonecpluspluscoder.com/2017/04/27/installing-java-8-jdk-os-x-using-homebrew/  
* http://www.lonecpluspluscoder.com/2017/10/08/installing-other-versions-of-the-java-jdk-via-homebrew/  

*********************************************************************

## install

```sh
# see which versions are available via brew
brew search java 
brew search openjdk

# install the version(s) you want
brew cask install java # latest version
brew cask install java8
brew cask install java11

# check your default version
java -version
which java
```

### java8

`brew install java8` is no longer available, but you can install Java 8 like this:

```sh
brew tap adoptopenjdk/openjdk
brew search adoptopenjdk
brew cask install adoptopenjdk8

# check it's installed
ls -1 /Library/Java/JavaVirtualMachines 

# add the java home directory
jenv add /Library/Java/JavaVirtualMachines/adoptopenjdk-8.jdk/Contents/Home/
```

*********************************************************************

## check versions

```sh
# list all currently installed versions
ls /Library/Java/JavaVirtualMachines/

# alternative way to find locations of previously installed Java JDK's installed at the default system locations
/usr/libexec/java_home -V # locations of all versions
/usr/libexec/java_home -v 1.8 # location of java8
/usr/libexec/java_home -v 11 # location of java11
echo $(/usr/libexec/java_home) # location of current system version
```

version numbers:

45 = Java 1.1  
46 = Java 1.2  
47 = Java 1.3  
48 = Java 1.4  
49 = Java 5  
50 = Java 6  
51 = Java 7  
52 = Java 8  
53 = Java 9  
54 = Java 10  
55 = Java 11  
56 = Java 12  
57 = Java 13  


*********************************************************************

## JAVA_HOME

One way to manage java versions is to manually set the `JAVA_HOME` environment variable which points to the file system location where the JDK or JRE was installed.

Add the following to `.zshrc` or `.bash_profile`:

```
export JAVA_8_HOME=$(/usr/libexec/java_home -v1.8)
export JAVA_11_HOME=$(/usr/libexec/java_home -v11)
alias java8='export JAVA_HOME=$JAVA_8_HOME'
alias java11='export JAVA_HOME=$JAVA_11_HOME'
```

Then simply use the relevant alias to switch version as needed.

**NB: For some reason setting JAVA_HOME doesn't work for me - but [jenv](#jenv) does.**

*********************************************************************

## jenv {#jenv}

> Java environment manager

* https://github.com/jenv/jenv  
* lets you switch between java versions  
* sets JAVA_HOME in a way that can be set globally, local to the current working directory or per shell  

```sh
brew install jenv
echo 'export PATH="$HOME/.jenv/bin:$PATH"' >> ~/.zshrc
echo 'eval "$(jenv init -)"' >> ~/.zshrc
source ~/.zshrc

# enable the export plugin to make sure JAVA_HOME is set 
jenv enable-plugin export
exec $SHELL -l # for me this runs source ~/.zshrc

# verify jenv was installed
jenv doctor

# add a specific version to the jenv list
jenv add $(/usr/libexec/java_home) # e.g. add current system version
jenv add $(/usr/libexec/java_home -v 1.8) # e.g. add java8
jenv add $(/usr/libexec/java_home -v 11) # e.g. add java11
jenv add $(/usr/libexec/java_home -v 14) # e.g. add java14

# see which java versions are now available to jenv
jenv versions
```

There are three options for setting versions: global, local and shell

global:

```sh
jenv global openjdk64-1.8.0.242
java -version
```

local:

```sh
# set a local version for the current working directory
# this will create a .java-version - jenv will load it when a shell is started from this directory
pwd
jenv local 11.0.2
exec $SHELL -l
cat .java-version
echo ${JAVA_HOME}
```

shell:

```sh
# set the version used in this particular shell session
jenv shell 11.0.2
```

*********************************************************************

## R

### rJava

* https://download.java.net/java/GA/jdk11/13/GPL/openjdk-11.0.1_osx-x64_bin.tar.gz

check that rJava loads properly - if it does, you have nothing to worry about :)

```{r, eval=FALSE}
library(rJava)
```

If you get an error message, try runnint this in a terminal (you many need to restart R afterwards)

```sh
R CMD javareconf 
```

### RStudio

For some later version of RStudio I get this error message when running `library(rJava)`:

```
Error: package or namespace load failed for ‘rJava’:
 .onLoad failed in loadNamespace() for 'rJava', details:
  call: dyn.load(file, DLLpath = DLLpath, ...)
  error: unable to load shared object '/Library/Frameworks/R.framework/Versions/3.6/Resources/library/rJava/libs/rJava.so':
  dlopen(/Library/Frameworks/R.framework/Versions/3.6/Resources/library/rJava/libs/rJava.so, 6): Library not loaded: /Library/Java/JavaVirtualMachines/jdk-11.0.1.jdk/Contents/Home/lib/server/libjvm.dylib
  Referenced from: /Library/Frameworks/R.framework/Versions/3.6/Resources/library/rJava/libs/rJava.so
  Reason: image not found
```

Running `R CMD javareconf` doesn't fix it and I also couldn't install `jdk-11.0.1`. 

However this hacky fix worked for me:

```sh
sudo mkdir -p /Library/Java/JavaVirtualMachines/jdk-11.0.1.jdk/Contents/Home/lib/server/
cd /Library/Java/JavaVirtualMachines/jdk-11.0.1.jdk/Contents/Home/lib/server/
sudo ln -s /Library/Java/JavaVirtualMachines/openjdk-11.0.2.jdk/Contents/Home/lib/server/libjvm.dylib libjvm.dylib
cd -
```

Version 1.2.1335 and earlier seem to work fine with Java 8 (e.g. version 1.8.0.192).

*********************************************************************

